\showyourwork: a workflow for open source scientific articles

### 2022-08-01 src/tex/ms.tex
more text

% Other imports
\usepackage{etoolbox}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{tikz}

@@

% Other stuff
\newcommand\syw{\texttt{showyourwork}\xspace}
\newcommand\fileurl[1]{https://github.com/showyourwork/showyourwork-paper/blob/\GitHubSHA/#1}

@@

\syw was designed to tackle these three issues, making it easy to develop, publish, and distribute truly open and reproducible research papers in astronomy and other scientific disciplines. 
At its core, it is a command-line tool that builds papers from a set of instructions contained in a \texttt{GitHub} repository and organized into \texttt{TeX} files, figure scripts, pipeline files, and configuration/specifications files.
Every time the user pushes a new commit to \texttt{GitHub}, the article is automatically built on the cloud using \texttt{GitHub Actions} and the resulting PDF is pushed to a separate branch of the repository. 
The build step---which sets up the conda environment, generates all figures from scratch (with intelligent caching), and compiles the PDF---acts as a unit test for the paper. 
If it passes, the paper is (by definition) reproducible.

@@

\section{Using {\protect\syw}}
\label{sec:usage}
In this section we provide basic, high-level instructions on how to install and use \syw.
These instructions are for version \texttt{0.3.0} of the code, the current version at the time of writing (July 2022) and are subject to change in future releases.
For more details, including custom commands, settings, examples, and troubleshooting tops, please refer to the documentation for the specific version of \syw at \url{https://show-your.work}.

@@

\noindent\begin{minipage}{\linewidth}
\begin{lstlisting}[
    style=bash
]
python -m pip install -U showyourwork
\end{lstlisting}
\end{minipage}

@@

\noindent\begin{minipage}{\linewidth}
\begin{lstlisting}[
    style=bash
]
git clone https://github.com/showyourwork/showyourwork
cd showyourwork
python -m pip install .
\end{lstlisting}
\end{minipage}

\subsection{Reproducing an article}
\label{sec:usage:reproduce}
Any project based on \syw can be reproduced by cloning its \texttt{GitHub} repository and running \syw. For example, to reproduce this paper, run:\\

\noindent\begin{minipage}{\linewidth}
\begin{lstlisting}[
    style=bash
]
git clone https://github.com/showyourwork/showyourwork-paper
cd showyourwork-paper
showyourwork
\end{lstlisting}
\end{minipage}

@@

\subsection{Repository structure}
\label{sec:usage:struct}
%
\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=0.8\linewidth]{figures/tree.pdf}
        \caption{
            The basic repository structure for an open source scientific article built with \syw.
            This figure was automatically generated from the \texttt{TikZ} code in \texttt{src/scripts/tree.tex} by specifying a custom command in \texttt{showyourwork.yml}.
        }
        \label{fig:tree}
        \script{tree.tex}
    \end{centering}
\end{figure}
%
Figure~\ref{fig:tree} shows the basic directory structure for a \syw article repository as of version \texttt{0.3.0}.
Repositories are comprised of a configuration files at the root level and a directory called \texttt{src} containing the manuscript files and all of the the code and workflow scripts needed to render the final article PDF.

@@

\begin{figure}[htp]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/eccentricity.pdf}
        \caption{
            The effect of binary eccentricity on the detectability of a \emph{LISA} gravitational wave source; reproduced from Figure 3 in \citet{Wagg2022}. 
            This figure was automatically generated from the script \texttt{src/scripts/eccentricity.py}.
        }
        \label{fig:eccentricity}
        \script{eccentricity.py}
    \end{centering}
\end{figure}

% \begin{figure}[htp]
%     \begin{centering}
%         \includegraphics[width=\linewidth]{figures/luhman16b.pdf}
%         \caption{
%             16 \emph{CRIRES} spectra of WISE 1049-5319B spanning a full rotation period of the brown dwarf; adapted from Figure 14 in \citet{Luger2021} and based on data from \citet{Crossfield2014}.
%             This figure was automatically generated from the script \texttt{src/scripts/luhman16b.py} and a dataset downloaded from \texttt{Zenodo}.
%         }
%         \label{fig:luhman16b}
%         \script{luhman16b.py}
%     \end{centering}
% \end{figure}

% \begin{figure}[htp]
%     \begin{centering}
%         \includegraphics[width=\linewidth]{figures/two_moons.pdf}
%         \caption{
%             A normalizing flow demonstrated on the two moons data set from \texttt{scikit-learn};
%             reproduced from Figure 1 in \citet{Crenshaw2022}.
%             This figure was automatically generated from the script \texttt{src/scripts/two\_moons.py}
%             and an intermediate dataset that was automatically cached on \texttt{Zenodo}.
%         }
%         \label{fig:two_moons}
%         \script{two_moons.py}
%     \end{centering}
% \end{figure}

@@

\begin{figure}[htp]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/dag.pdf}
        \caption{
            A directed acyclic graph (DAG) showing the complete list of dependencies for the article. 
            Scripts are shown in green, \texttt{Zenodo} deposits in blue, datasets in purple, and \texttt{TeX} files in red.
            This figure is located in the \texttt{src/static} directory and, unlike the other figures in this article, is version controlled by \texttt{git}. The \texttt{src/static} directory is reserved for figures that are not programmatically generated and simply get copied over to the output directory at compile time.
        }
        \label{fig:dag}
    \end{centering}
\end{figure}

>>>>

% Other imports
\usepackage{etoolbox}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage[figuresright]{rotating}

@@

% Other stuff
\newcommand\syw{\texttt{showyourwork}\xspace}
\newcommand\repourl{https://github.com/showyourwork/showyourwork-paper}
\newcommand\commiturl{\repourl/tree/\GitHubSHA}
\newcommand\fileurl[1]{\repourl/blob/\GitHubSHA/#1}

@@

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/overview.pdf}
        \caption{
            Overview of the \syw workflow for a scientific article.
            All code and instructions to compile the article exist within a GitHub repository (top center), optionally synced with an Overleaf project (top right).
            Upon every commit, a GitHub Action is triggered to build the article in an isolated environment.
            Scripts are executed to generate figures and other workflow output (bottom center), optionally downloading version-controlled datasets from Zenodo (top left), and optionally caching intermediate results on Zenodo Sandbox (bottom left).
            Finally, the compiled article PDF, alongside an arXiv-compatible tarball, are generated and pushed to a special branch on the GitHub repository.
        }
        \label{fig:overview}
    \end{centering}
\end{figure}

\syw was designed to tackle these three issues, making it easy for authors to develop, publish, and distribute truly open and reproducible research papers in astronomy and other scientific disciplines. 
At its core, it is a command-line tool that builds papers from a set of instructions contained in a \texttt{GitHub} repository and organized into \texttt{TeX} files, figure scripts, pipeline files, and configuration/specifications files (see Figure~\ref{fig:overview}).
Every time the user pushes a new commit to \texttt{GitHub}, the article is automatically built on the cloud using \texttt{GitHub Actions} and the resulting PDF is pushed to a separate branch of the repository. 
The build step---which sets up the conda environment, generates all figures from scratch (with intelligent caching), and compiles the PDF---acts as a unit test for the paper. 
If it passes, the paper is (by definition) reproducible.

@@

\section{Using {\protect\syw}}
\label{sec:usage}
In this section we provide basic, high-level instructions on how to install and use \syw.
These instructions are for version \texttt{0.3.0} of the code, the current version at the time of writing (July 2022) and are subject to change in future releases.
For more details, including custom commands, settings, examples, and troubleshooting tips, please refer to the documentation for the specific version of \syw at \url{https://show-your.work}.

@@

\noindent\begin{minipage}{\linewidth}
\begin{lstlisting}[
    style=bash
]
pip install -U showyourwork
\end{lstlisting}
\end{minipage}

@@

\noindent\begin{minipage}{\linewidth}
\begin{lstlisting}[
    style=bash
]
git clone https://github.com/showyourwork/showyourwork
cd showyourwork
python setup.py install .
\end{lstlisting}
\end{minipage}

\subsection{Reproducing an article}
\label{sec:usage:reproduce}
Any project based on \syw can be reproduced by cloning its \texttt{GitHub} repository and running \syw. For example, to reproduce the exact version of this paper that you are currently reading, run:\\

\noindent\begin{minipage}{\linewidth}
\begin{lstlisting}[
    style=bash,
    escapeinside={|}{|}
]
git clone https://github.com/showyourwork/showyourwork-paper
cd showyourwork-paper
git checkout |\GitHubSHA|
showyourwork
\end{lstlisting}
\end{minipage}

@@

\subsection{Repository structure}
\label{sec:usage:struct}
%
\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=0.8\linewidth]{figures/tree.pdf}
        \caption{
            The basic repository structure for an open source scientific article built with \syw.
            The article you are reading was generated from a repository with this exact structure; you can check it out \href{\commiturl}{here}.
            This figure was automatically generated from the \texttt{TikZ} code in \texttt{src/scripts/tree.tex} by specifying a custom command in \texttt{showyourwork.yml}.
        }
        \label{fig:tree}
        \script{tree.tex}
    \end{centering}
\end{figure}
%
Figure~\ref{fig:tree} shows the basic directory structure for a \syw article repository as of version \texttt{0.3.0}.
Repositories are comprised of a configuration files at the root level and a directory called \texttt{src} containing the manuscript files and all of the the code and workflow scripts needed to render the final article PDF.

@@

To establish dependencies between scripts and figures included in the TeX manuscript, users may provide the custom \texttt{\textbackslash script\{{\color{lsthilite}script}\}} command within the \texttt{figure} environment, where \texttt{\color{lsthilite}script} is the name of the script \syw should execute to generate the figure(s) within the current environment.
Additional dependencies may be specified explicitly in the \texttt{showyourwork.yml} configuration file or implicitly via custom rules in the \texttt{Snakefile}.

As a seconde example, consider Figure~\ref{fig:luhman16b}, which we reproduce from \citet{Luger2021}. 
Unlike Figure~\ref{fig:eccentricity}, this figure has an external dependency: a Zenodo-hosted dataset containing the \emph{CRIRES} observations of the spectrum of a brown dwarf.
Users simply specify the DOI of the dataset and the files they wish to make dependencies of a given figure in \texttt{showyourwork.yml}, and \syw will download the data as needed.

Finally, Figure~\ref{fig:two_moons} is an example of a figure depending on an intermediate result that is cached on Zenodo Sandbox.

\clearpage

\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=0.75\linewidth]{figures/eccentricity.pdf}
        \caption{
            The effect of binary eccentricity on the detectability of a \emph{LISA} gravitational wave source; reproduced from Figure 3 in \citet{Wagg2022}. 
            This figure was automatically generated from the script \texttt{src/scripts/eccentricity.py}.
            The GitHub icon in the margin is a clickable link pointing to the exact version of the script on GitHub used to produce this figure.
        }
        \label{fig:eccentricity}
        \script{eccentricity.py}
    \end{centering}
\end{figure}

\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=0.75\linewidth]{figures/luhman16b.pdf}
        \caption{
            16 \emph{CRIRES} spectra of WISE 1049-5319B spanning a full rotation period of the brown dwarf; adapted from Figure 14 in \citet{Luger2021} and based on data from \citet{Crossfield2014}.
            This figure was automatically generated from the script \texttt{src/scripts/luhman16b.py} and a dataset downloaded from \texttt{Zenodo}.
            In addition to the GitHub icon linking to the script, this figure also has a dataset icon linking to the Zenodo deposit that hosts the data needed to generate it.
        }
        \label{fig:luhman16b}
        \script{luhman16b.py}
    \end{centering}
\end{figure}

\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=0.75\linewidth]{figures/two_moons.pdf}
        \caption{
            A normalizing flow demonstrated on the two moons data set from \texttt{scikit-learn};
            reproduced from Figure 1 in \citet{Crenshaw2022}.
            This figure was automatically generated from the script \texttt{src/scripts/two\_moons.py}
            and an intermediate dataset that was automatically cached on \texttt{Zenodo}.
        }
        \label{fig:two_moons}
        \script{two_moons.py}
    \end{centering}
\end{figure}

@@

\begin{sidewaysfigure}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/dag.pdf}
        \caption{
            A directed acyclic graph (DAG) showing the complete list of dependencies for the article. 
            Scripts are shown in green, \texttt{Zenodo} deposits in blue, datasets in purple, and \texttt{TeX} files in red.
            This figure is located in the \texttt{src/static} directory and, unlike the other figures in this article, is version controlled by \texttt{git}. The \texttt{src/static} directory is reserved for figures that are not programmatically generated and simply get copied over to the output directory at compile time.
        }
        \label{fig:dag}
    \end{centering}
\end{sidewaysfigure}
